Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-09 bbbq_article.tex
Remove evolutionary conservation

Evolutionary conservation scores are shown in 
figures \ref{fig:evolutionary_conservation_8k} (topology),
\ref{fig:evolutionary_conservation_8k_violin} (per 
protein with at least 1 TMH, violin)
and \ref{fig:evolutionary_conservation_8k_merged} (grouped 
from all proteins that have at least 1 TMH).
See figures
\ref{fig:evolutionary_conservation_8k_all} and 
\ref{fig:evolutionary_conservation_8k_all_violin}
for the scores for all proteins).

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores.png}
  \caption{
    Conservation scores for the aligned SARS-CoV-2 proteomes,
    for proteins that have at least one TMH,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
    The (non-logarithmic) scale of the conservation score has been removed
    for readability.
    Conservation scores do not start at residue 1 when
    an aligned protein sequence outflanks the consensus sequence 
    at the N-terminal
  }
  \label{fig:evolutionary_conservation_8k}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_merged_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    with conservation scores group from all proteins that have at least one TMH,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
  }
  \label{fig:evolutionary_conservation_8k_merged}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    for proteins that have at least one TMH,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
  }
  \label{fig:evolutionary_conservation_8k_violin}
\end{figure}

@@

We have removed the 25 selenoproteins found in the human
reference proteome, as both prediction tools struggle with 
selenocysteine, the 21st amino acid incorporated in human proteins.
TMHMM gives a clear error message, forcing us to remove
these selenoproteins from the analysis.
How PureseqTM deals with selenocysteine is unclear,
as it gives no error, nor is it mentioned anywhere,
nor do the authors respond to emails. Due to this, we removed the
selenoproteins from our analysis.
Another bug in PureseqTM is that it does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1_HUMAN protein, short for 'T cell receptor delta diversity 1',
is, however, two peptides long. We manually added the prediction
that both peptides reside in the cytosol.

@@

@@

Per target proteome (i.e. human, SARS-CoV-2, Mycobacterium tuberculosis),
we collected all 9-mers (for MHC-I) and 15-mers (for MHC-II),
after removing the selenoproteins and proteins that are shorter
than the epitope length.
From these epitopes, per MHC haplotype,
we predicted the IC50 using \verb;epitope-prediction; (for MHC-I)
and MHCnuggets (for MHC-II). 
Here, we show the IC50 value per haplotype that
is used to determine if a peptide binds to the haplotype's MHC.

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc1_2_normalized.png}
  \caption{
    Normalized proportion of MHC-I epitopes overlapping with TMHs
    for human, viral and bacterial proteomes.
    Legend: covid = SARS-CoV-2, flua = influenza A, hepa = hepatitis A,
    human = homo sapiens, myco = Mycobacterium tuberculosis, 
    polio = poliovirus, rhino = rhinovirus
  }
  \label{fig:f_tmh_mhc1_normalized}
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc2_2_normalized.png}
  \caption{
    Normalized proportion of MHC-II epitopes overlapping with TMHs
    for human, viral and bacterial proteomes.
    Legend: covid = SARS-CoV-2, flua = influenza A, hepa = hepatitis A,
    human = homo sapiens, myco = Mycobacterium tuberculosis, 
    polio = poliovirus, rhino = rhinovirus
  }
  \label{fig:f_tmh_mhc2_normalized}
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores.png}
  \caption{
    Conservation scores for the aligned SARS-CoV-2 proteomes,
    for proteins that have at least one TMH,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
    The (non-logarithmic) scale of the conservation score has been removed
    for readability.
    Conservation scores do not start at residue 1 when
    an aligned protein sequence outflanks the consensus sequence 
    at the N-terminal
  }
  \label{fig:evolutionary_conservation_8k}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_merged_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    with conservation scores group from all proteins that have at least one TMH,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
  }
  \label{fig:evolutionary_conservation_8k_merged}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    for proteins that have at least one TMH,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
  }
  \label{fig:evolutionary_conservation_8k_violin}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/8000_42/scores_all.png}
  \caption{
    Conservation scores for the aligned SARS-CoV-2 proteomes,
    for all proteins,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
    The (non-logarithmic) scale of the conservation score has been removed
    for readability.
    Conservation scores do not start at residue 1 when
    an aligned protein sequence outflanks the consensus sequence 
    at the N-terminal.
  }
  \label{fig:evolutionary_conservation_8k_all}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin_all.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    for all proteins,
    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
  }
  \label{fig:evolutionary_conservation_8k_all_violin}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{All unique sequences}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
  \caption{
    Conservation scores for the aligned SARS-CoV-2 proteomes,
    for proteins that have at least one TMH,
    using all unique sequences \richel{which I doubt is valid}.
    The (non-logarithmic) scale of the conservation score has been removed
    for readability.
    Conservation scores do not start at residue 1 when
    an aligned protein sequence outflanks the consensus sequence 
    at the N-terminal
  }
  \label{fig:evolutionary_conservation_all_unique}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    with conservation scores group from all proteins that have at least one TMH,
    using all unique sequences \richel{which I doubt is valid}.
  }
  \label{fig:evolutionary_conservation_all_unique_merged}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    for proteins that have at least one TMH,
    using all unique sequences \richel{which I doubt is valid}.
  }
  \label{fig:evolutionary_conservation_all_unique_violin}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
  \caption{
    Conservation scores for the aligned SARS-CoV-2 proteomes,
    for proteins that have at least one TMH,
    using all unique sequences \richel{which I doubt is valid}.
    The (non-logarithmic) scale of the conservation score has been removed
    for readability.
    Conservation scores do not start at residue 1 when
    an aligned protein sequence outflanks the consensus sequence 
    at the N-terminal
  }
  \label{fig:evolutionary_conservation_all_unique}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    with conservation scores group from all proteins that have at least one TMH,
    using all unique sequences \richel{which I doubt is valid}.
  }
  \label{fig:evolutionary_conservation_all_unique_merged}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    for proteins that have at least one TMH,
    using all unique sequences \richel{which I doubt is valid}.
  }
  \label{fig:evolutionary_conservation_all_unique_violin}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/scores_all.png}
  \caption{
    Conservation scores for the aligned SARS-CoV-2 proteomes,
    for all proteins,
    using all unique sequences \richel{which I doubt is valid}.
    The (non-logarithmic) scale of the conservation score has been removed
    for readability.
    Conservation scores do not start at residue 1 when
    an aligned protein sequence outflanks the consensus sequence 
    at the N-terminal.
  }
  \label{fig:evolutionary_conservation_all_unique_all}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_2/scores_violin_all.png}
  \caption{
    Evolutionary conservation scores of TMH versus non-TMH
    residues in the SARS-CoV-2 proteome,
    for all proteins,
    using all unique sequences \richel{which I doubt is valid}.
  }
  \label{fig:evolutionary_conservation_all_unique_all_violin}
\end{figure}

>>>>

@@

We have removed the 25 selenoproteins found in the human
reference proteome, as both prediction tools struggle with 
selenocysteine, the 21st amino acid incorporated in human proteins.
TMHMM gives a clear error message, forcing us to remove
these selenoproteins from the analysis.
How PureseqTM deals with selenocysteine is unclear,
as it gives no error, nor is it mentioned anywhere,
nor do the authors respond to emails. Due to this, we removed the
selenoproteins from our analysis.
Another bug in PureseqTM is that it does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1\_HUMAN protein, short for 'T cell receptor delta diversity 1',
is, however, two peptides long. We manually added the prediction
that both peptides reside in the cytosol.

@@

Table \ref{tab:uniprot_ids} shows an overview of the Uniprot IDs
for the reference proteomes used. Note that for Homo sapiens,
we did not use the full reference proteome, but a representative
and smaller reference proteome instead.

@@

Per target proteome (i.e. human, SARS-CoV-2, Mycobacterium tuberculosis),
we collected all 9-mers (for MHC-I) and 15-mers (for MHC-II),
after removing the selenoproteins and proteins that are shorter
than the epitope length.
From these epitopes, per MHC haplotype,
we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
and MHCnuggets (for MHC-II). 
Here, we show the IC50 value per haplotype that
is used to determine if a peptide binds to the haplotype's MHC
for MHC-I (see \ref{tab:ic50_binders_mhc1}) and 
MHC-II (see \ref{tab:ic50_binders_mhc2}).

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc1_2_normalized.png}
  \caption{
    Normalized proportion of MHC-I epitopes overlapping with TMHs
    for human, viral and bacterial proteomes.
    Legend: covid = SARS-CoV-2,
    human = homo sapiens, myco = Mycobacterium tuberculosis
  }
  \label{fig:f_tmh_mhc1_normalized}
\end{figure}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc2_2_normalized.png}
  \caption{
    Normalized proportion of MHC-II epitopes overlapping with TMHs
    for human, viral and bacterial proteomes.
    Legend: covid = SARS-CoV-2,
    human = homo sapiens, myco = Mycobacterium tuberculosis
  }
  \label{fig:f_tmh_mhc2_normalized}
\end{figure}

@@
